---
title: "Statistical Analyses"
output:
  html_document:
    df_print: paged
---

# Preparation
## Set up 
```{r setup}
getwd()
knitr::opts_knit$set(root.dir = "/home/Maren.Krafft/R_docs/Working_files")
```

## Load libraries
```{r lib, include=FALSE}
library("data.table")
library(ggplot2)
library(tidyr)
library(plyr) 
library(dplyr)
library(furniture)
library(purrr)   
library(dplyr)      
library(foreign)   
library(effsize)
```

```{r}
#install.packages("glmbb")
library(broom)
#devtools::install_github("atyre2/NRES803")
#pak::pkg_install("atyre2/NRES803")
library("NRES803")
library("radiant.data")
```
## Load dataset
```{r Load, echo=TRUE}
#load("final/dat_excl.Rda")
dat_excl2 <- dat_excl
#dat_clean2 <- dat_clean
```
## Creating work data sets
```{r define}
dat <- dat_excl2
dat2 <- dat_excl2[heart_rate>= 60 & heart_rate <=100,]
summary(dat)
```

```{r function, include=FALSE}
  hist.default <- function(x,
                         breaks = "Scott",
                         freq = NULL,
                         include.lowest = TRUE,
                         normalcurve = TRUE,
                         right = TRUE,
                         density = NULL,
                         angle = 45,
                         col = NULL,
                         border = NULL,
                         main = paste("Histogram of", xname),
                         ylim = NULL,
                         xlab = xname,
                         ylab = NULL,
                         axes = TRUE,
                         plot = TRUE,
                         labels = FALSE,
                         warn.unused = TRUE,
                         ...)  {

  # https://stackoverflow.com/a/20078645/4575331
  xname <- paste(deparse(substitute(x), 500), collapse = "\n")

  suppressWarnings(
    h <- graphics::hist.default(
      x = x,
      breaks = breaks,
      freq = freq,
      include.lowest = include.lowest,
      right = right,
      density = density,
      angle = angle,
      col = col,
      border = border,
      main = main,
      ylim = ylim,
      xlab = xlab,
      ylab = ylab,
      axes = axes,
      plot = plot,
      labels = labels,
      warn.unused = warn.unused,
      ...
    )
  )

  if (normalcurve == TRUE & plot == TRUE) {
    x <- x[!is.na(x)]
    xfit <- seq(min(x), max(x), length = 40)
    yfit <- dnorm(xfit, mean = mean(x), sd = sd(x))
    if (isTRUE(freq) | (is.null(freq) & is.null(density))) {
      yfit <- yfit * diff(h$mids[1:2]) * length(x)
    }
    lines(xfit, yfit, col = "black", lwd = 2)
  }

  if (plot == TRUE) {
    invisible(h)
  } else {
    h
  }
  }
  
```

## Summary statistics
```{r summary}
#summary(dat)

dat %>%
  group_by(variant) %>%
  table1(age, sex, ca, tsh, ill_resp, ill_heart, ill_arrythmia, champers_pq,  p_duration, qrs_duration, qt, QTc_B, QTc_F, heart_rate, rr,  Q_duration_V3,  R_duration_V3,  S_duration_V3,  T_duration_V3)

dat %>%
  table1(age, sex, ca, tsh, ill_resp, ill_heart, ill_arrythmia, champers_pq,  p_duration, qrs_duration, qt, QTc_B, QTc_F, heart_rate, rr,  Q_duration_V3,  R_duration_V3,  S_duration_V3,  T_duration_V3)


dat2 %>%
  group_by(variant) %>%
  table1(age, sex, ca, tsh, ill_resp, ill_heart, ill_arrythmia, champers_pq,  p_duration, qrs_duration, qt, QTc_B, QTc_F, heart_rate, rr,  Q_duration_V3,  R_duration_V3,  S_duration_V3,  T_duration_V3)

dat2 %>%
  group_by(variant) %>%
  table1(age, sex, ca, tsh, ill_resp, ill_heart, ill_arrythmia, champers_pq,  p_duration, qrs_duration, qt, QTc_B, QTc_F, heart_rate, rr,  Q_duration_V3,  R_duration_V3,  S_duration_V3,  T_duration_V3)

dat2 %>%
  table1(age, sex, ca, tsh, ill_resp, ill_heart, ill_arrythmia, champers_pq,  p_duration, qrs_duration, qt, QTc_B, QTc_F, heart_rate, rr,  Q_duration_V3,  R_duration_V3,  S_duration_V3,  T_duration_V3)

dat2 %>%
  group_by(variant) %>%
  table1(bmi, fit_ind)

dat2 %>%
  group_by(sex, variant) %>%
  table1(QTc_B, QTc_F)
```

```{r}
mean(dat$QTc_B[dat$variant ==1])
mean(dat$QTc_B[dat$variant ==0])
```

# Statistical Analyses

## Welch T-Test: Complete dataset
```{r}
# Test for normally 
shapiro.test(dat$QTc_B)
# Test for equal variance
fligner.test(QTc_B ~ variant, dat)

# Welch Two Sample t-test
t.test(QTc_B ~ variant, data = dat, detailed = true)

cohen.d(QTc_B ~ variant, var.equal = FALSE, data = dat)

#Framingham
# Test for normally 
shapiro.test(dat$QTc_F)
# Test for equal variance
fligner.test(QTc_F ~ variant, dat)

# Welch Two Sample t-test
t.test(QTc_F ~ variant, data = dat, detailed = true)

cohen.d(QTc_F ~ variant, var.equal = FALSE, data = dat)
```

## Welch T-Test. Dataset reduced to individuals with a normal heart rate
```{r}
#Bazett
# Test for normally 
shapiro.test(dat2$QTc_B)
# Test for equal variance
fligner.test(QTc_B ~ variant, dat2)

# Welch Two Sample t-test
t.test(QTc_B ~ variant, data = dat2, detailed = true)

cohen.d(QTc_B ~ variant, var.equal = FALSE, data = dat2)


#Framingham

# Test for normally 
shapiro.test(dat2$QTc_F)
# Test for equal variance
fligner.test(QTc_F ~ variant, dat2)

# Welch Two Sample t-test
t.test(QTc_F ~ variant, data = dat2, detailed = true)

cohen.d(QTc_F ~ variant, var.equal = FALSE, data = dat2)
```


## Correlation for covariables

```{r}
dat_test <-dat[,c("QTc_B","age","ca","tsh", "syst_bp", "ill_resp", "ill_heart", "ill_arrythmia", "c1", "c2", "c3", "c4", "fit_ind", "bmi", "med_other")]
sapply(dat_test, class)
cor(dat_test, use="complete.obs")
#cor(dat[,c(age,sex,ca,tsh,diast_bp,syst_bp,ill_resp,ill_heart,ill_arrythmia,c1,c2,c3,c4)])

boxplot(dat$QTc_B)
pairs(dat_test[,1:5], pch = 19, lower.panel = NULL)
pairs(dat_test[,c(1, 6:10)], pch = 19, lower.panel = NULL)
pairs(dat_test[,c(1, 11:15)], pch = 19, lower.panel = NULL)
```


## Linear Regression QT interval corrected with Bazett

### Complete dataset
```{r bazett dat comp}

dat_ana <- dat
a <-lm(formula = QTc_B ~ variant +age+ sex+ age*sex + c(age*age) ,data=dat_ana)
summary(a)

b <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat_ana)
summary(b)

c <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi ,data=dat_ana)
summary(c)

d <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + fit_ind,data=dat_ana)
summary(d)

e <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind,data=dat_ana)
summary(e)

f <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4+ med_other,data=dat_ana)
summary(f)

g <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind + med_other,data=dat_ana)
summary(g)

```
```{r}
check_assumptions(a)
check_assumptions(b)
check_assumptions(c)
check_assumptions(d)
check_assumptions(e)
check_assumptions(f)
check_assumptions(g)
```

### Reduced to individuals with a heart rate in the normal range
```{r bazett dat red}

dat_ana <- dat2
a <-lm(formula = QTc_B ~ variant +age+ sex+ age*sex + c(age*age) ,data=dat_ana)
summary(a)

b <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat_ana)
summary(b)

c <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi ,data=dat_ana)
summary(c)

d <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + fit_ind,data=dat_ana)
summary(d)

e<-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind,data=dat_ana)
summary(r)

f <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + med_other,data=dat_ana)
summary(f)

g <-lm(formula = QTc_B ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind + med_other,data=dat_ana)
summary(g)

```

```{r}
check_assumptions(a)
check_assumptions(b)
check_assumptions(c)
check_assumptions(d)
check_assumptions(e)
check_assumptions(f)
check_assumptions(g)
```

## Linear Regression QT interval corrected with Framingham 
### Complete dataset
```{r framingham dat comp}

dat_ana <- dat
a <-lm(formula = QTc_F ~ variant +age+ sex+ age*sex + c(age*age) ,data=dat_ana)
summary(a)

b <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat_ana)
summary(b)

c <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi ,data=dat_ana)
summary(c)

d <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + fit_ind,data=dat_ana)
summary(d)

e <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind,data=dat_ana)
summary(e)

f <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + med_other,data=dat_ana)
summary(f)

g <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind  + med_other,data=dat_ana)
summary(g)

```
```{r}
check_assumptions(a)
check_assumptions(b)
check_assumptions(c)
check_assumptions(d)
check_assumptions(e)
check_assumptions(f)
check_assumptions(g)
```


### Reduced to individuals with a heart rate in the normal range
```{r framingham dat red}

dat_ana <- dat2
a <-lm(formula = QTc_F ~ variant +age+ sex+ age*sex + c(age*age) ,data=dat_ana)
summary(a)

b <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat_ana)
summary(b)

c <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi ,data=dat_ana)
summary(c)

d <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + fit_ind,data=dat_ana)
summary(d)

e <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind,data=dat_ana)
summary(e)

f <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + med_other,data=dat_ana)
summary(f)

g <-lm(formula = QTc_F ~  variant + age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 + bmi + fit_ind + med_other,data=dat_ana)
summary(g)

```

```{r}
check_assumptions(a)
check_assumptions(b)
check_assumptions(c)
check_assumptions(d)
check_assumptions(e)
check_assumptions(f)
check_assumptions(g)
```

### Linear regression for every covariate individually
```{r bazett single}

m <-lm(formula = QTc_B ~ variant ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~  age ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~  sex ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ as.numeric(fit_ind),data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ bmi ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~  ca ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ tsh ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ diast_bp ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ syst_bp ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ ill_resp ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ ill_heart,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ ill_arrythmia ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ med_psych  ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ med_presc ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ med_other ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ c1 ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ c2 ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ c3 ,data=dat2)
summary(m)
check_assumptions(m)

m <-lm(formula = QTc_B ~ c4 ,data=dat2)
summary(m)
check_assumptions(m)
```

# Excursion other ECG timings
## T-Test ECG Timings
```{r}
# Test for normally 
shapiro.test(dat$heart_rate)
# Test for equal variance
fligner.test(heart_rate ~ variant, dat)
# Welch Two Sample t-test
t.test(heart_rate ~ variant, data = dat, detailed = true)
cohen.d(heart_rate ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$heart_rate,method =  "spearman")


# Test for normally 
shapiro.test(dat$p_duration)
# Test for equal variance
fligner.test(p_duration ~ variant, dat)
# Welch Two Sample t-test
t.test(p_duration ~ variant, data = dat, detailed = true)
cohen.d(p_duration ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$p_duration,method =  "spearman")

# Test for normally 
shapiro.test(dat$champers_pq)
# Test for equal variance
fligner.test(champers_pq ~ variant, dat)
# Welch Two Sample t-test
t.test(champers_pq ~ variant, data = dat, detailed = true)
cohen.d(champers_pq ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$champers_pq,method =  "spearman")


# Test for normally 
shapiro.test(dat$Q_duration_V3)
# Test for equal variance
fligner.test(Q_duration_V3 ~ variant, dat)
# Welch Two Sample t-test
t.test(Q_duration_V3 ~ variant, data = dat, detailed = true)
cohen.d(Q_duration_V3 ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$Q_duration_V3,method =  "spearman")


# Test for normally 
shapiro.test(dat$R_duration_V3)
# Test for equal variance
fligner.test(R_duration_V3 ~ variant, dat)
# Welch Two Sample t-test
t.test(R_duration_V3 ~ variant, data = dat, detailed = true)
cohen.d(R_duration_V3 ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$R_duration_V3,method =  "spearman")


# Test for normally 
shapiro.test(dat$S_duration_V3)
# Test for equal variance
fligner.test(S_duration_V3 ~ variant, dat)
# Welch Two Sample t-test
t.test(S_duration_V3 ~ variant, data = dat, detailed = true)
cohen.d(S_duration_V3 ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$S_duration_V3,method =  "spearman")


# Test for normally 
shapiro.test(dat$qrs_duration)
# Test for equal variance
fligner.test(qrs_duration ~ variant, dat)
# Welch Two Sample t-test
t.test(qrs_duration ~ variant, data = dat, detailed = true)
cohen.d(qrs_duration ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$qrs_duration,method =  "spearman")


# Test for normally 
shapiro.test(dat$T_duration_V3)
# Test for equal variance
fligner.test(T_duration_V3 ~ variant, dat)
# Welch Two Sample t-test
t.test(T_duration_V3 ~ variant, data = dat, detailed = true)
cohen.d(T_duration_V3 ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$T_duration_V3,method =  "spearman")


# Test for normally 
shapiro.test(dat$rr)
# Test for equal variance
fligner.test(rr ~ variant, dat)
# Welch Two Sample t-test
t.test(rr ~ variant, data = dat, detailed = true)
cohen.d(rr ~ variant, var.equal = FALSE, data = dat)
cor.test ( as.numeric(dat$variant), dat$rr,method =  "spearman")

```

## Linear regression for individual ECG timings

```{r other}

b <-lm(formula = heart_rate ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = p_duration ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = champers_pq ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = Q_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = R_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = S_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = qrs_duration ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = T_duration_V3 ~ variant+ age+ bmi + sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)

b <-lm(formula = rr ~ variant+ age+ bmi + sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat)
summary(b)
check_assumptions(b)
```


```{r other}

b <-lm(formula = heart_rate ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = p_duration ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = champers_pq ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = Q_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = R_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = S_duration_V3 ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = qrs_duration ~ variant+ age+ sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = T_duration_V3 ~ variant+ age+ bmi + sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)

b <-lm(formula = rr ~ variant+ age+ bmi + sex+ age*sex + c(age*age) + ca +tsh+ diast_bp +syst_bp+ ill_resp +ill_heart +ill_arrythmia  +c1 +c2+c3+c4 ,data=dat2)
summary(b)
check_assumptions(b)
```


## Analyses regarding QT time in normal range
```{r}
summary(dat[variant == 0 & sex== "male" &  QTc_B > 450, variant])
summary(dat[variant == 1& sex== "male" &  QTc_B > 450, variant])
summary(dat[variant == 0&sex== "female" &  QTc_B > 460, variant])
summary(dat[variant == 1&sex== "female" &  QTc_B > 460, variant])


summary(dat[variant == 0 & sex== "male" &  QTc_F > 450, variant])
summary(dat[variant == 1& sex== "male" &  QTc_F > 450, variant])
summary(dat[variant == 0&sex== "female" &  QTc_F > 460, variant])
summary(dat[variant == 1&sex== "female" &  QTc_F > 460, variant])


summary(dat2[variant == 0 & sex== "male" &  QTc_B > 450, variant])
summary(dat2[variant == 1& sex== "male" &  QTc_B > 450, variant])
summary(dat2[variant == 0&sex== "female" &  QTc_B > 460, variant])
summary(dat2[variant == 1&sex== "female" &  QTc_B > 460, variant])


summary(dat2[variant == 0 & sex== "male" &  QTc_F > 450, variant])
summary(dat2[variant == 1& sex== "male" &  QTc_F > 450, variant])
summary(dat2[variant == 0&sex== "female" &  QTc_F > 460, variant])
summary(dat2[variant == 1&sex== "female" &  QTc_F > 460, variant])
```


```{r}
dat_all2 <- dat
dat_all2$toohigh_B <- ifelse(dat_all2$sex == "female", dat_all2$QTc_B >460, dat_all2$QTc_B >450)
dat_all2$toohigh_F <- ifelse(dat_all2$sex == "female", dat_all2$QTc_F >460, dat_all2$QTc_F >450)

chisq <- chisq.test( dat_all4$toohigh_B,dat_all4$variant,correct = FALSE)
chisq

dat_all4 <- dat2
dat_all4$toohigh_B <- ifelse(dat_all4$sex == "female", dat_all4$QTc_B >460, dat_all4$QTc_B >450)
dat_all4$toohigh_F <- ifelse(dat_all4$sex == "female", dat_all4$QTc_F >460, dat_all4$QTc_F >450)

chisq <- chisq.test( dat_all2$toohigh_B,dat_all2$variant,correct = FALSE)
chisq

```
