---
title: "PheWAS"
output: html_document
date: "2023-01-14"
---

## Reading in the genotypes
```{r include=FALSE}
library(data.table)
library(R.utils)
library(ggplot2)
library(survminer)
library(survival)
#install.packages("devtools")
#install.packages(c("dplyr","tidyr","ggplot2","MASS","meta","ggrepel","DT"))
#devtools::install_github("PheWAS/PheWAS")
library(PheWAS)
library(tibble)
library(stringr)
library(data.table)
library(plyr)
library(tidyr)
library(glue)
library(dplyr)
```

### Exomedaten
```{r exom, eval=FALSE, include=FALSE}
d <- read.table("IDs.txt",header = TRUE, sep = "")
d$X3.38613787.G.A_A <- as.factor(d$X3.38613787.G.A_A)
d$X3.38613787.G.A_HET <- as.factor(d$X3.38613787.G.A_HET)
d$SEX <- as.factor(d$SEX)
genotype <- d[, c(5,3,2)]
colnames(genotype) <- c("ID", "genotype", "sex")

save(genotype,file="genotype_exom.Rda")
summary(genotype)
#summary(d)
```
### Imputation
```{r microarray}
dat <- read.table("Variant",header = TRUE, sep = "")
colnames(dat) <- c("eid_old", "eid", "sex", "genotype")
dat$sex <- as.factor(dat$sex)
dat$genotype <- as.factor(dat$genotype)

summary(dat)
save(dat,file="genotype_micro.Rda")
```
### Genotypen vergleichen
```{r compare_files, eval=FALSE, include=FALSE}
dat[dat$eid_old == 1151210,]

compare <- dat[dat$eid %in% genotype$ID,]
merged <- merge(genotype, dat, by.x ="ID", by.y = "eid")
a <- ifelse(merged$genotype.x==merged$genotype.y,0,1)
subset(merged, merged$genotype.x != merged$genotype.y)

d[d$IID == 1151210,]
```

# Column overview csv phenotype file
```{r getalldata, echo=FALSE}
data_overview <- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", nrows = 2)
```


```{r}

last_follow <- which( colnames(data_overview)=="191-0.0" )
b <- grep("21003", colnames(data_overview))
test <- c(1,  b)

tested<- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", select = test)
last_assessment <-tested
save(last_assessment,file="last_assessment.Rda")
```
### Get latest assessment age
```{r}
assess <- melt(last_assessment[1:502415,], id = "eid")
assess <- assess[!is.na(assess$value), c(1,3)]
assess <- setorder(assess, cols = "eid")   
head(assess)

latest_assessment <- aggregate(cbind(latest_followup_age = value) ~ eid, assess, max)
save(latest_assessment,file="latest_assessment.Rda")
```

```{r}
total_with_lastdiag <- merge(total_with_lastdiag, latest_assessment, by.x= "eid", by.y = "eid", all.x=TRUE )

count(total_with_lastdiag$age < total_with_lastdiag$latest_followup_age)
total_with_lastdiag$last_age <- ifelse(total_with_lastdiag$latest_followup_age > total_with_lastdiag$age | is.na(total_with_lastdiag$age), total_with_lastdiag$latest_followup_age, total_with_lastdiag$age)
```


### Get general information 
```{r generalcolumns, echo=FALSE}
#Columnes for ICD-Codes
death_age <- grep("40007", colnames(data_overview))
death_age <- death_age[1]
birth_year <- which( colnames(data_overview)=="34-0.0" )
sex  <- which(colnames(data_overview)=="31-0.0" )
pc <- grep("22009", colnames(data_overview))
pc <- pc[1:10]
birth_month<- which( colnames(data_overview)=="52-0.0" )
ethnic <- grep("22006", colnames(data_overview))

general <- c(1, sex, birth_year, birth_month, death_age, pc, ethnic)

general <- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", select = general)

total <- merge(dat, general, by = "eid")

colnames(total) <- c("eid", "eid_old", "sex", "genotype", "sex_new", "birth_year", "birth_month" , "death_age_1", paste0("pc", 1:10), "white_british")

total <- total[total$eid != 1410515 & total$eid != 3938173 ,]
total <- data.table(total)
total$birthdate <- paste(total$birth_year, ifelse(nchar(total$birth_month) ==2, total$birth_month, paste("0", total$birth_month, sep="")), "01", sep = "-")
total <-total[,-21]
save(total,file="total.Rda")

```

### Extract ICD COdes
```{r}
icd1 <- grep("41270", colnames(data_overview))
icd1 <- c(1, icd1)

icds <- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", select = icd1)
save(icds,file="icds.Rda")
```

### Melt the ICD-codes to bring it in the form eid -icd
```{r}
all_icd_1 <- melt(icds[1:100000,], id = "eid")
all_icd_1 <- all_icd_1[!(is.na(all_icd_1$value) | all_icd_1$value==""), c(1,3)]
all_icd_1 <- setorder(all_icd_1, cols = "eid")   

all_icd_2 <- melt(icds[100001:200000,], id = "eid")
all_icd_2 <- all_icd_2[!(is.na(all_icd_2$value) | all_icd_2$value==""), c(1,3)]
all_icd_2 <- setorder(all_icd_2, cols = "eid")   

all_icd_3 <- melt(icds[200001:300000,], id = "eid")
all_icd_3 <- all_icd_3[!(is.na(all_icd_3$value) | all_icd_3$value==""), c(1,3)]
all_icd_3 <- setorder(all_icd_3, cols = "eid")   

all_icd_4 <- melt(icds[300001:400000,], id = "eid")
all_icd_4 <- all_icd_4[!(is.na(all_icd_4$value) | all_icd_4$value==""), c(1,3)]
all_icd_4 <- setorder(all_icd_4, cols = "eid")   

all_icd_5 <- melt(icds[400001:502415,], id = "eid")
all_icd_5 <- all_icd_5[!(is.na(all_icd_5$value)) & all_icd_5$value!="", c(1,3)]
all_icd_5 <- setorder(all_icd_5, cols = "eid")   

all_icds <- rbind(all_icd_1, all_icd_2, all_icd_3, all_icd_4, all_icd_5)
#save(all_icds,file="all_icds.Rda")
```

### Extract diagnosis dates 
```{r}
dates <- grep("41280", colnames(data_overview))
dates <- c(1, dates)

diag_dates <- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", select = dates)

save(diag_dates,file="date_columns.Rda")
```

### Melt the dates to bring it in the form eid - date
```{r}
all_1 <- melt(diag_dates[1:100000,], id = "eid")
all_1 <- all_1[!is.na(all_1$value), c(1,3)]
all_1 <- setorder(all_1, cols = "eid")   

all_2 <- melt(diag_dates[100001:200000,], id = "eid")
all_2 <- all_2[!(is.na(all_2$value) ), c(1,3)]
all_2 <- setorder(all_2, cols = "eid")   

all_3 <- melt(diag_dates[200001:300000,], id = "eid")
all_3 <- all_3[!(is.na(all_3$value) ), c(1,3)]
all_3 <- setorder(all_3, cols = "eid")   

all_4 <- melt(diag_dates[300001:400000,], id = "eid")
all_4 <- all_4[!(is.na(all_4$value)), c(1,3)]
all_4 <- setorder(all_4, cols = "eid")   

all_5 <- melt(diag_dates[400001:502415,], id = "eid")
all_5 <- all_5[!(is.na(all_5$value)), c(1,3)]
all_5 <- setorder(all_5, cols = "eid")   

all_dates <- rbind(all_1, all_2, all_3, all_4, all_5)
save(all_dates,file="all_dates.Rda")
#summary(all_dates)
```

### Get last diagnosis and add it to the total datatable 
```{r}
birthdate <- total[, c("eid", "birthdate")]
dates_merged <- merge(all_dates, birthdate, by = "eid")

dates_merged$diff_age <- as.numeric(difftime(as.Date(dates_merged$value), as.Date(dates_merged$birthdate), "days"))/365

save(dates_merged,file="dates_merged.Rda")

results <- aggregate(cbind(date_last_diag = diff_age) ~ eid, dates_merged, max)

total <- merge(total, results, by = "eid",all = TRUE)

save(total,file="total_with_lastdiag.Rda")
```

### Extracts Death Diagnoses
```{r}
death_diag <- grep("40001", colnames(data_overview))
death_diag_sec <- grep("40002", colnames(data_overview))

death <- c(1, death_diag, death_diag_sec)

diag_death <- fread("~/ukbiobank/ukbiobank_77717/original/phenotypes/ukb49791.csv", select = death)

#save(diag_death,file="diag_death.Rda")
```

### Melt the death diagnoses to bring it in the form eid - death diagnosis
```{r}
all_death_1 <- melt(diag_death[1:100000,], id = "eid")
all_death_1 <- all_death_1[!(is.na(all_death_1$value) | all_death_1$value==""), c(1,3)]
all_death_1 <- setorder(all_death_1, cols = "eid")   

all_death_2 <- melt(diag_death[100001:200000,], id = "eid")
all_death_2 <- all_death_2[!(is.na(all_death_2$value) | all_death_2$value==""), c(1,3)]
all_death_2 <- setorder(all_death_2, cols = "eid")   

all_death_3 <- melt(diag_death[200001:300000,], id = "eid")
all_death_3 <- all_death_3[!(is.na(all_death_3$value) | all_death_3$value==""), c(1,3)]
all_death_3 <- setorder(all_death_3, cols = "eid")   

all_death_4 <- melt(diag_death[300001:400000,], id = "eid")
all_death_4 <- all_death_4[!(is.na(all_death_4$value) | all_death_4$value==""), c(1,3)]
all_death_4 <- setorder(all_death_4, cols = "eid")   

all_death_5 <- melt(diag_death[400001:502415,], id = "eid")
all_death_5 <- all_death_5[!(is.na(all_death_5$value)) & all_death_5$value!="", c(1,3)]
all_death_5 <- setorder(all_death_5, cols = "eid")   
#head(all_death_5)

all_death_diag_death <- rbind(all_death_1, all_death_2, all_death_3, all_death_4, all_death_5)

all_death_diag <- all_death_diag_death
save(all_death_diag,file="all_death.Rda")
```

### Create reduced ICD-Code tabke
```{r}
all_icds_incldeath <- rbind(all_icds, all_death_diag)
names(all_icds_incldeath)[names(all_icds_incldeath) == 'value'] <- 'icd'
save(all_icds_incldeath,file="all_icds_incldeath.Rda")

all_icd_red <- all_icds_incldeath
all_icd_red$icd <-substring(all_icd_red$value,1,3)
all_icd_red <- all_icd_red[,-2]

all_icd_red <- all_icd_red[!duplicated(all_icd_red), ]
save(all_icd_red, file="all_icd_red.Rda")
```

### Create final total dataset
```{r}
load("total_with_lastdiag.Rda")

dat_total <- total
names(dat_total)[names(dat_total) == 'death_age_1'] <- 'death_age'

dat_total$has_death_diag <- ifelse(dat_total$eid %in% all_death_diag$eid, TRUE, FALSE)
dat_total$age <- round(ifelse(is.na(dat_total$death_age), dat_total$date_last_diag, dat_total$death_age), 2)

dat_total$age_sq <- dat_total$age^2

dat_total$sex_age <- dat_total$age*as.numeric(dat_total$sex)

dat_total <- dat_total[,-c(2,5,19)]

save(dat_total,file="Phewas/total_complet.Rda")
write.csv(dat_total, "Phewas/total_complete.csv", row.names = FALSE)

dat_wNA <- dat_total
dat_total<- dat_total[!is.na(dat_total$age)]

save(dat_total,file="withoutAge/total_complete_woNa.Rda")
write.csv(dat_total, "withoutAge/total_complete_woNA.csv", row.names = FALSE)
```

### Prepare ICD-Code and merge with general information file
```{r}

#load("withoutAge/total_complete.Rda")
load("all_icds_incldeath.Rda")
load("all_icd_red.Rda")

# For reduced icd 
I4<- all_icd_red %>% filter(str_detect(icd,"^I4"))
I66<- all_icd_red %>% filter(str_detect(icd,"^I66"))
I5<- all_icd_red %>% filter(str_detect(icd,"^I5"))

Phenotypes<-rbind(I4, I66,I5)
ICDs_red<- unique(Phenotypes$icd)
save(ICDs_red,file="withoutAge/icds_red.Rda")

total2 <- dat_total
for (ICD in ICDs_red){
  ICDlist<-unique(Phenotypes[icd==ICD]$eid)
  total2[[ICD]]<-ifelse(total2$eid%in%ICDlist,TRUE,FALSE)
}

total_red <- total2
save(total_red,file="withoutAge/total_red.Rda")

#For detailed icd
I4<- all_icds_incldeath %>% filter(str_detect(icd,"^I4"))
I66<- all_icds_incldeath %>% filter(str_detect(icd,"^I66"))
I5<- all_icds_incldeath %>% filter(str_detect(icd,"^I5"))

Phenotypes<-rbind(I4, I66, I5)
ICDs<- unique(Phenotypes$icd)
save(ICDs,file="withoutAge/icds_all.Rda")

total2 <- dat_total

for (ICD in ICDs){
  ICDlist<-unique(Phenotypes[icd==ICD]$eid)
  total2[[ICD]]<-ifelse(total2$eid%in%ICDlist,TRUE,FALSE)
}

total_all <- total2
save(total_all,file="withoutAge/total_all.Rda")
```


### Phewas
```{r}
sub <- c("eid", "sex", paste0("pc",1:10),"age", "age_sq", "sex_age")

results_het_all=phewas_ext(data= total_all, phenotypes=ICDs,genotypes="genotype", covariates= sub[-1], additive.genotypes=T, cores=1) 
save(results_het_all,file="withoutAge/phewas_het_all_new.Rda")
write.csv(results_het_all,file="withoutAge/phewas_het_all_new.csv", row.names = FALSE)

results_het_red=phewas_ext(data= total_red, phenotypes=ICDs_red,genotypes="genotype", covariates= sub[-1], additive.genotypes=T, cores=1)
save(results_het_red,file="withoutAge/phewas_het_red_new.Rda")
write.csv(results_het_red,file="withoutAge/phewas_het_red_new.csv", row.names = FALSE)
```

# Add ICD description
```{r}
ICD_names <- read.table("ICD_Names_comp",header = TRUE, sep = "\t")
setorder(results_het_all, cols =  "phenotype")  
setorder(results_het_red, cols =  "phenotype")  
results_het_all <-merge(ICD_names,results_het_all,by="phenotype", all.y=TRUE)
results_het_red <-merge(ICD_names,results_het_red,by="phenotype", all.y=TRUE)
```

# Sums frequency
```{r}
#sum <- read.csv("sumgen.csv")

gen_1_icd_red <- total_red[total_red$genotype ==1, -c(1:22)]
a <- colSums(gen_1_icd_red == TRUE)
dt_red = data.table( icd = colnames(gen_1_icd_red), genTRUE = a)

gen_1_icd_all <- total_all[total_all$genotype ==1, -c(1:22)]
b <- colSums(gen_1_icd_all == TRUE)
dt_all = data.table( icd = colnames(gen_1_icd_all), genTRUE = b)

results_het_all <- merge( results_het_all,dt_all, by.x = "phenotype", by.y = "icd")

results_het_red <- merge( results_het_red,dt_red, by.x = "phenotype", by.y = "icd")

save(results_het_all,file="withoutAge/results_het_all_sum.Rda")
write.csv(results_het_all,file="withoutAge/results_het_all_sum.csv", row.names = FALSE)

save(results_het_red,file="withoutAge/results_het_red_sum.Rda")
write.csv(results_het_red,file="withoutAge/results_het_red_sum.csv", row.names = FALSE)
```

### Round p-values and exclude I48 in specified ICD-Codes

```{r}
results_red <- data.table(results_het_red)
results_all <- data.table(results_het_all)
results_all <- results_all[phenotype != "I48",]

results_red$p_short <-  round(results_red$p, 2)
results_all$p_short <-  round(results_all$p, 2)
results_red[results_red$phenotype == "I48","p_short"] <- round(results_red[results_red$phenotype == "I48",p],5)
results_red[results_red$phenotype == "I40","p_short"] <- round(results_red[results_red$phenotype == "I40",p],3)


results_all[phenotype == "I409","p_short"] <- round(results_all[phenotype == "I409",p],3)
results_all[phenotype == "I441","p_short"] <- round(results_all[phenotype == "I441",p],3)
results_all[phenotype == "I455","p_short"] <- round(results_all[phenotype == "I455",p],4)
results_all[phenotype == "I480","p_short"] <- round(results_all[phenotype == "I480",p],3)


save(results_all,file="withoutAge/results_all_phewas.Rda")
save(results_red,file="withoutAge/results_red_phewas.Rda")
write.csv(results_all, "withoutAge/results_all.csv", row.names = FALSE)
write.csv(results_red, "withoutAge/results_red.csv", row.names = FALSE)
```




### Add Confidence interval 

```{r}
results_all$CIpos <- exp(results_all$beta+1.96*results_all$SE)
results_all$CIneg <- exp(results_all$beta-1.96*results_all$SE)

results_red$CIpos <- exp(results_red$beta+1.96*results_red$SE)
results_red$CIneg <- exp(results_red$beta-1.96*results_red$SE)
```

```{r}
results_all <- results_all[,-c(3,9,13:15,17:18)]
results_red <- results_red[,-c(3,9,13:15,17:18) ]

save(results_all,file="RDAs/results_all_phewas2.Rda")
save(results_red,file="RDAs/results_red_phewas2.Rda")
write.csv(results_all, "CSV/results_all2.csv", row.names = FALSE)
write.csv(results_red, "CSV/results_red2.csv", row.names = FALSE)
stri_sub(results_all$phenotype, 4, 3) <- "."
#ab hier ausführen

results_all <- results_all[genTRUE >= 4, ]
results_red <- results_red[genTRUE >= 4, ]
save(results_all,file="RDAs/results_all.Rda")
save(results_red,file="RDAs/results_red.Rda")
write.csv(results_all, "CSV/results_all.csv", row.names = FALSE)
write.csv(results_red, "CSV/results_red.csv", row.names = FALSE)
```

```{r}

p.adjust(results_red$p, method = "fdr")

```

### ggplot libraries
```{r}
library(ggplot2)
library(ggrepel)
library(scales)
library(stringi)
```

### ggplot Theme
```{r}

theme_rbook <- function(base_size = 8, base_family = "", base_line_size = base_size/22, 
                        base_rect_size = base_size/22) {         
  theme( 
    axis.title = element_text(size = 10, margin = margin(0,0, 20 ,0)),                               
    axis.text.y = element_text(color = "black",size = 8, margin = margin(0,0, 0 ,0)),    
    axis.text.x = element_text(color = "black",angle=45, hjust=1, size=8),
    axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
    axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)),
    plot.caption = element_text(size = 10, face = "italic"),  
    plot.title = element_text(size = 10, 
                                  hjust = 0.5, 
                                   face = "bold", 
                                 margin = margin(10,0,10,0)),
    panel.background = element_rect(fill="white"),
    panel.grid.major = element_line(colour = "grey", size = 0.2),
    axis.line = element_line(size = 1, colour = "black"),
    strip.background =element_rect(fill = "#cddcdd"),
    panel.border = element_rect(colour = "black", fill=NA, size=0.2),
    strip.text = element_text(colour = "black"),
    legend.key=element_blank(),legend.margin=margin(0,0,0,3),
    legend.box.spacing = unit(10, "pt"),
    legend.position="right",
    legend.background=element_blank(),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 6)
    
    
    
  )
}
```


```{r}
wild = "#1A5276"
het = "#FF8300"
```

### Phewas Plots
```{r}
plot_icd_red<-ggplot(results_red, aes(x=paste(phenotype,"/ ",desc), y=OR,label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point( color = wild) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", color =wild) +
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(size = 2.5, nudge_x=.2, direction = "y") +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  theme(plot.margin = margin(0,0,0, 1.5, "cm"))+
  ggtitle("Association between the genotype and ICD categories") +
  theme(plot.margin = margin(0,0,0, 3, "cm"))
plot_icd_red
ggsave("PLOT/plot_icd_red_new2.pdf",width = 8, height = 8)
ggsave("PLOT/plot_icd_red_new2_s.png",width = 10, height = 7)

outliershape <- ifelse(results_all$OR<2,16,17)

#results_all$icd_correct <- stri_sub(results_all$phenotype, 4, 3) <- "."
plot_icd_all<-ggplot(results_all, aes(x=paste(phenotype,"/ ",desc), y=ifelse(OR<2,OR,2),label=p_short, 3))+
  geom_hline(aes(yintercept= 1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(shape =outliershape, color =wild) +
  geom_errorbar(aes(ymin=CIneg, ymax=ifelse(CIpos> 2.5, 2.5, CIpos)), width=0, alpha=0.7, stat = "identity", color = wild) +
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(size = 2.5, nudge_x=.3, direction = "y") +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Association between the genotype and ICD subcategories") +
  theme(plot.margin = margin(0,0,0, 3, "cm"))
 # scale_shape_manual(values = c(1, 4))
plot_icd_all
ggsave("PLOT/plot_icd_all_new2.pdf", height = 6 , width = 7)
ggsave("PLOT/plot_icd_all_new2_s.png",width =  10, height = 6)
```

# Change color!!!
```{r}
library(survival)
library(survminer)

#load(all_death.Rda)
I48<- all_death_diag %>% filter(str_detect(value,"^I48"))

ICDs<- unique(I48$value)

surv_all <- dat_total
surv_all$sss <- surv_all$eid%in%I48$eid #Todesdiagnose
survobj <- survfit(Surv(age, sss) ~ genotype, data = surv_all)

out3 <- ggsurvplot(
  survobj,
  fun = "event",
  censor = F, 
  
  ggtheme = theme_rbook(),
  #theme(legend.position = "left"),
  legend.title = "Genotype",
  legend.labs = c("Wildtype", "Heterozygous"),
  size = 0.4,
  palette = c(wild, het),

  grid = NULL,
  conf.int = TRUE,

 xlim = c(25, 85),
  break.x.by= 5
 #break.y.by= 0.01
 ) +ylab("Probability to die from atrial fibrillation")+xlab("Age (years)") 

out3
ggsave("PLOT/Survival_death.pdf", height = 8 , width = 8)
ggsave("PLOT/Survival_death.png", height = 4 , width = 4)
```

### Proportion without atrial fibrilation
```{r}
#load(date_columns.Rda)
#load(.Rda)

#Get IDs with atrial fibrillation and flutter
I48<- all_icds_incldeath %>% filter(str_detect(icd,"^I48"))
eids <- unique(I48$eid)

#Get every individuals row that has this diagnosis
diag_dates_I48 <- diag_dates[eid %in% eids,]
diag_icds_I48 <- icds[eid %in% eids,]
#summary(diag_dates_I48)

#Change the structure
diag_dates_melted <- melt(diag_dates_I48, id = "eid")
diag_icds_melted <- melt(diag_icds_I48, id = "eid")

#Merge dates and diagnosis
diag_icd_combined <- cbind(diag_dates_melted, diag_icds_melted)
diag_icd_combined <- diag_icd_combined[,-c(2,5)]
colnames(diag_icd_combined) <- c("eid", "date","eid2", "icd")
#save(diag_icd_combined, file="diag_icd_combined.Rda")



I48_combined <- diag_icd_combined %>% filter(str_detect(icd,"^I48"))

#save(I48_combined, file="I48_combined.Rda")


birthdate <- dat_total[,c("eid", "birthdate")]

dates_merged <- merge(I48_combined, birthdate, by = "eid", all.x = TRUE)

dates_merged$diff_age <- round(as.numeric(difftime(as.Date(dates_merged$date), as.Date(dates_merged$birthdate), "days"))/365,2)

results <- aggregate(cbind(first_diag_I48 = diff_age) ~ eid, dates_merged, min)


I48_firstdiag <- results
#save(I48_firstdiag, file="I48_firstdiag.Rda")


#get death diags with I48
death_diag_merged <- merge(all_death_diag, dat_total[,c("eid","death_age")], by = "eid", all.x = TRUE)
I48_death <- death_diag_merged %>% filter(str_detect(value,"^I48"))
I48_death <- I48_death[!is.na(I48_death$death_age),] #56 NAs
colnames(I48_death) <-c("eid", "icd", "first_diag_I48")
I48_firstdiag_incl_death <- rbind(I48_firstdiag, I48_death[,c(1,3)])

I48_first <- aggregate(cbind(first_diag_I48 = first_diag_I48) ~ eid, I48_firstdiag_incl_death, min)
save(I48_first, file="withoutAge/I48_first.Rda")

```

```{r}
dat_total_I48 <- merge(dat_total, I48_first, by = "eid", all.x =TRUE)


save(dat_total_I48, file="dat_total_I48_.Rda")


```


```{r}
firstdiag <- ggplot(dat_total_I48, aes(x=genotype, y=first_diag_I48, fill=genotype)) + 
  geom_violin(width = 0.4, lwd = 0.3) + 
  geom_boxplot(outlier.size=0.2,width=0.2, color="black", alpha=0.2, lwd = 0.3) +
  scale_fill_manual(values=c(wild, het)) +
  theme_rbook() +
  ylab("Age at first diagnosis of I48") +
  
   scale_x_discrete(name ="Genotype", labels=c("0" = "Wildtype", "1" = "Heterozygous"))+
  labs( title ="Age comparison of first atrial fibrillation diagnosis")+
  theme(legend.position = "none")
firstdiag
ggsave("PLOT/First_diagnosis_I48.pdf", height = 8 , width = 8)
ggsave("PLOT/First_diagnosis_I48.png", height = 4 , width = 4)

mean(dat_total_I48$first_diag_I48[dat_total_I48$genotype ==1], na.rm=TRUE)
mean(dat_total_I48$first_diag_I48[dat_total_I48$genotype ==0], na.rm=TRUE)
sd(dat_total_I48$first_diag_I48[dat_total_I48$genotype ==1], na.rm=TRUE)
sd(dat_total_I48$first_diag_I48[dat_total_I48$genotype ==0], na.rm=TRUE)

model <- lm(first_diag_I48 ~ genotype + sex + age + birth_year + age_sq + sex_age, data = dat_total_I48)
summary(model)

```
Survplot Proportion without atrial fibrillation

```{r}
surv_all_1 <- data.table(dat_total_I48)
surv_all_1[, rel_age := first_diag_I48]
surv_all_1[, hasdiag := ifelse(is.na(first_diag_I48), F, T)]
surv_all_1[is.na(rel_age), rel_age := age]
survobj2 <- survfit(Surv(rel_age, hasdiag) ~ genotype, data = surv_all_1)

out4 <- ggsurvplot(
  survobj2,
  #fun = "event",
  censor = F, 
  
  ggtheme = theme_rbook(),
  #theme(legend.position = "left"),
  legend.title = "Genotype",
  legend.labs = c("Wildtype", "Heterozygous"),
  size = 0.4,
  palette = c(wild, het),

  grid = NULL,
  conf.int = TRUE,
 # Add risk table
 xlim = c(25, 83),fontsize = 3,
  break.x.by= 5,
 legend = "none"
# break.y.by= 0.01
  
) +ylab("Proportion without atrial fibrillation and flutter")+xlab("Age (years)") 
out4
ggsave("PLOT/Without_I48.pdf", height = 8 , width = 8)
ggsave("PLOT/Without_I48.png", height = 4 , width = 4)
```



### Meta-Analysis

```{r}
phewas_finn2 <- read.table("phewas_finn_all.txt",header = TRUE, sep = ",")
phewas_finn2 <- phewas_finn2[phewas_finn2$genotype != "hom",]
ICD_names <- read.table("ICD_Names_comp",header = TRUE, sep = "\t")
phewas_finn2 <-merge(ICD_names,phewas_finn2,by="phenotype", all.y=TRUE)

save(phewas_finn2, file="RDAs/phewas_finn_before_cleaning.Rda")
phewas_finn2 <- phewas_finn2[,-c(3,9,13:18,21)]

phewas_finn2$Database <- "FinnGen"

phewas_finn_all <- data.table(phewas_finn2[phewas_finn2$red == 0, ])
phewas_finn_red <- data.table(phewas_finn2[phewas_finn2$red == 1, ])
stri_sub(phewas_finn_all$phenotype, 4, 3) <- "."



setorder(phewas_finn_all, cols = "phenotype")
setorder(phewas_finn_red, cols = "phenotype")


phewas_finn_red[, p_short:= ifelse(p > 0.05, round(p, 2),0)]

phewas_finn_red[phenotype == "I42","p_short"] <- round(phewas_finn_red[phenotype == "I42",p],3)
phewas_finn_red[phenotype == "I47","p_short"] <- round(phewas_finn_red[phenotype == "I47",p],7)
phewas_finn_red[phenotype == "I48","p_short"] <- round(phewas_finn_red[phenotype == "I48",p],26)
phewas_finn_red[phenotype == "I49","p_short"] <- round(phewas_finn_red[phenotype == "I49",p],5)
phewas_finn_red[phenotype == "I50","p_short"] <- round(phewas_finn_red[phenotype == "I50",p],4)

phewas_finn_all[, p_short:= ifelse(p > 0.05, round(p, 2),0)]

phewas_finn_all[phenotype == "I44.1","p_short"] <- round(phewas_finn_all[phenotype == "I44.1",p],3)
phewas_finn_all[phenotype == "I44.7","p_short"] <- round(phewas_finn_all[phenotype == "I44.7",p],3)
phewas_finn_all[phenotype == "I47.1","p_short"] <- round(phewas_finn_all[phenotype == "I47.1",p],6)
phewas_finn_all[phenotype == "I48.0","p_short"] <- round(phewas_finn_all[phenotype == "I48.0",p],8)
phewas_finn_all[phenotype == "I48.9","p_short"] <- round(phewas_finn_all[phenotype == "I48.9",p],9)
phewas_finn_all[phenotype == "I49.5","p_short"] <- round(phewas_finn_all[phenotype == "I49.5",p],20)
phewas_finn_all[phenotype == "I49.9","p_short"] <- round(phewas_finn_all[phenotype == "I49.9",p],6)
phewas_finn_all[phenotype == "I50.1","p_short"] <- round(phewas_finn_all[phenotype == "I50.1",p],4)
phewas_finn_all[phenotype == "I50.9","p_short"] <- round(phewas_finn_all[phenotype == "I50.9",p],4)


phewas_finn_red$CIpos <- exp(phewas_finn_red$beta+1.96*phewas_finn_red$SE)
phewas_finn_red$CIneg <- exp(phewas_finn_red$beta-1.96*phewas_finn_red$SE)

phewas_finn_all$CIpos <- exp(phewas_finn_all$beta+1.96*phewas_finn_all$SE)
phewas_finn_all$CIneg <- exp(phewas_finn_all$beta-1.96*phewas_finn_all$SE)

write.csv(phewas_finn_all, "CSV/phewas_finn_all_all.csv", row.names = FALSE)
write.csv(phewas_finn_red, "CSV/phewas_finn_red_all.csv", row.names = FALSE)


phewas_finn_all <- phewas_finn_all[phenotype %in% results_all$phenotype,]

phewas_finn_red <- phewas_finn_red[phenotype %in% results_red$phenotype,]

save(phewas_finn_red,file="RDAs/phewas_finn_red.Rda")
save(phewas_finn_all,file="RDAs/phewas_finn_all.Rda")
write.csv(phewas_finn_all, "CSV/phewas_finn_all.csv", row.names = FALSE)
write.csv(phewas_finn_red, "CSV/phewas_finn_red.csv", row.names = FALSE)
```


```{r}
plot_finn_red<-ggplot(phewas_finn_red, aes(x=paste(phenotype,"/ ",desc), y=OR,label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point( color = wild) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", color = wild) +
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(size = 2.5, nudge_x=.3, direction = "y") +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Association between the variant and ICD categories - FinnGen") 
plot_finn_red
ggsave("PLOT/plot_phewas_finngen_red.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_finngen_red.png", height = 4 , width = 8)

plot_finn_all<-ggplot(phewas_finn_all, aes(x=paste(phenotype,"/ ",desc), y=OR,label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(color = wild) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", color =wild) +
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(size = 2.5, nudge_x=.3, direction = "y") +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2.5)) +
  xlab("ICD-Codes and Descriptions") +
  ylab("Odds Ratio") +
  ggtitle("Association between the variant and ICD subcategories - FinnGen") +
  theme(plot.margin = margin(0,0,0, 0.5, "cm"))
plot_finn_all
ggsave("PLOT/plot_phewas_finngen_all.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_finngen_all.png", height = 4 , width = 8)
```

```{r}
results_all <- results_all[genTRUE >= 4, ]
results_red <- results_red[genTRUE >= 4, ]

results_red$Database <- "UK Biobank"
complete_phewas_red <- data.table(dplyr::bind_rows(results_red, phewas_finn_red[phenotype %in% results_red$phenotype]))

results_all$Database <- "UK Biobank"
complete_phewas_all <- dplyr::bind_rows(results_all, phewas_finn_all[phenotype %in% results_all$phenotype])
setorder(complete_phewas_all, cols = "phenotype")

save(complete_phewas_red,file="RDAs/complete_phewas_red.Rda")
save(complete_phewas_all,file="RDAs/complete_phewas_all.Rda")
write.csv(complete_phewas_red,file="CSV/complete_phewas_red.csv", row.names = FALSE)
write.csv(complete_phewas_all,file="CSV/complete_phewas_all.csv", row.names = FALSE)
```


```{r}


plot_phewas_red<-ggplot(complete_phewas_red[phenotype !="I51"], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(nudge_x = .3,
    direction = "y",
    size = 2.5) +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Association comparison between Finngen and UK Biobank - ICD categories") +
  scale_color_manual(values = c(wild, het))+
  scale_fill_discrete("Database", labels = c("FinnGen", "UK Biobank")) +
  theme(legend.position = "right")
  
plot_phewas_red
#ggsave("PLOT/plot_phewas_meta.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_meta_red.png", height = 4 , width = 8)

plot_phewas_all<-ggplot(complete_phewas_all[-c(29:31)], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  #theme(legend.position="none") +
  geom_text_repel(nudge_x = .5,
    direction = "y",
    size = 2.5) +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2.5)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Comparison between Finngen and UK Biobank - ICD subcategories") +
  scale_color_manual(values = c(wild, het)) +
  theme(plot.margin = margin(0,0,0, 0.5, "cm"))
  
plot_phewas_all
#ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.png", height = 4 , width = 8)

```
```{r}


plot_phewas_red<-ggplot(complete_phewas_red[phenotype !="I51"], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  geom_text_repel(nudge_x = .3,
    direction = "y",
    size = 2.5) +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Association comparison between Finngen and UK Biobank - ICD categories") +
  scale_color_manual(values = c(wild, het))+
  scale_fill_discrete("Database", labels = c("FinnGen", "UK Biobank")) +
  theme(legend.position = "right")
  
plot_phewas_red
#ggsave("PLOT/plot_phewas_meta.pdf", height = 8 , width = 8)
#ggsave("PLOT/plot_phewas_meta_red.png", height = 4 , width = 8)

plot_phewas_all<-ggplot(complete_phewas_all[-c(29:31)], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  #theme(legend.position="none") +
  geom_text_repel(nudge_x = .5,
    direction = "y",
    size = 2.5) +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(0, 2.5)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Comparison between Finngen and UK Biobank - ICD subcategories") +
  scale_color_manual(values = c(wild, het)) +
  theme(plot.margin = margin(0,0,0, 1.5, "cm"))
  
plot_phewas_all
#ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.pdf", height = 8 , width = 8)
#ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.png", height = 4 , width = 8)

```


```{r}


plot_phewas_red<-ggplot(complete_phewas_red[phenotype %in% c("I47", "I48", "I49", "I50")], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  theme(legend.position="none") +
  #geom_text(position = position_dodge(width = 1), size = 3)+
  geom_text_repel(nudge_x = .3,direction = "y",size = 2.5, segment.color = 'transparent') +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(-0.5, 2.5)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Association comparison between Finngen and UK Biobank - ICD categories") +
  scale_color_manual(values = c(wild, het))+
  scale_fill_discrete("Database", labels = c("FinnGen", "UK Biobank")) +
  theme(legend.position = "right")  +
  theme(plot.margin = margin(0,0,0, 1.5, "cm"))
  
plot_phewas_red
#ggsave("PLOT/plot_phewas_meta.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_meta_red_onlyfinnrel.png", height = 6 , width = 7)

plot_phewas_all<-ggplot(complete_phewas_all[phenotype %in% c("I47.1", "I48.0", "I48.9", "I49.5", "I49.9")], aes(x=paste(phenotype,"/ ",desc), y=OR, group = Database, col = Database, label=p_short))+
  geom_hline(aes(yintercept=1), linetype="dashed", linewidth=0.3, alpha=0.5)+
  geom_point(position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin=CIneg, ymax=CIpos), width=0, alpha=0.7, stat = "identity", position = position_dodge(0.3))+
  theme_rbook() +  #geom_text(hjust=0, vjust=0) +
  #theme(legend.position="none") +
  #geom_text( size = 2.5)+
  geom_text_repel(nudge_x = .3,direction = "y",size = 2.5, segment.color = 'transparent') +
  scale_x_discrete(labels = label_wrap(40)) +
  coord_cartesian(ylim = c(-0.5, 2.5)) +
  xlab("ICD-codes and descriptions") +
  ylab("Odds ratio") +
  ggtitle("Comparison between Finngen and UK Biobank - ICD subcategories") +
  scale_color_manual(values = c(wild, het)) +
  theme(plot.margin = margin(0,0,0, 1.5, "cm"))
  
plot_phewas_all
#ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.pdf", height = 8 , width = 8)
ggsave("PLOT/plot_phewas_meta_all_new_onlyfinnp.png", height = 6 , width = 7)

```
```{r}
sub_red <-complete_phewas_red[phenotype %in% c("I47", "I48", "I49", "I50") & Database == "UK Biobank"]
p.adjust(sub_red$p, method = "fdr")
sub_all <-complete_phewas_all[phenotype %in% c("I47.1", "I48.0", "I48.9", "I49.5", "I49.9") & Database == "UK Biobank"]
p.adjust(sub_all$p, method = "fdr")
```


```{r}
library(psych)
library(vtable)
a <-describeBy(dat_total[,c(4,6,18,20,21,22)], dat_total$genotype, mat = TRUE) 
write.csv(a, "CSV/summary_stats.csv", row.names = FALSE)
summary(dat_total)
a <- 
  st(dat_total, group = 'genotype')
write.csv(a, "CSV/summarystatistics.csv", row.names = FALSE)

```

```{r}
library(furniture)

dat_total %>%
  group_by(genotype) %>%
  table1(age, sex)
3*(mean(dat_total$age)- median(dat_total$age))/ sd(dat_total$age)

dat_total %>%
  group_by(genotype ) %>%
  table1(age, sex, death_age)
```


Missing: Descriptive plots age sex genotype and distributions

```{r age, fig.width=4, fig.height=4}
#axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"), 
ybreaks = seq(0,500000,2000)
bw = 1
n_obs = sum(!is.na(dat_total$age))

ggb <- ggplot(dat_total, aes(age)) +
  geom_histogram(aes(y = ..density..), binwidth = bw, na.rm=TRUE, colour = wild, fill = "white") +
  theme_rbook() +
  stat_function(fun = dnorm, args = list(mean = mean(dat_total$age), sd = sd(dat_total$age))) +
  scale_y_continuous("Counts", breaks = round(ybreaks / (bw * n_obs),3), labels = ybreaks) +
  scale_x_continuous(breaks = seq(0, 100, 5)) +
  labs(title="Age distribution",x="Age (Years)", y = "Count")
ggb
ggsave(paste("PLOT/age.pdf", sep = ""), height = 4 , width = 4)
ggsave(paste("PLOT/age.png", sep = ""), height = 4 , width = 4)

ggb <- ggplot(data = dat_total, mapping = aes(x = genotype, y =age, fill =genotype)) + 
    geom_boxplot(outlier.size=0.2,na.rm=TRUE, colour = "black")+
  theme_rbook() +
  labs(x = "Genotype", y= "Age (years)", title ="Age comparison")  + 
  theme(legend.position = "none")+
  scale_x_discrete(labels=c("0" = "Wildtype", "1" = "Heterozygous")) +
  scale_fill_manual(values=c("0"=wild,"1"=het))
  #stat_summary(fun="mean")
ggb
ggsave(paste("PLOT/ageboxpl.pdf", sep = ""), height = 4, width = 4)
ggsave(paste("PLOT/ageboxpl.png", sep = ""), height = 4, width = 4)
```

```{r sex, fig.width=6, fig.height=6}

ggb <- ggplot(dat_total, aes(sex, age, fill = genotype)) +
  geom_boxplot(outlier.size=0.2, na.rm=TRUE, colour = "black") +
  theme_rbook() +
  labs(title="Sex-age relation",x="Sex", y = "Age (years)") +
  scale_fill_manual(values=c("0"=wild,"1"=het),name = "Genotype", labels=c("0" = "Wildtype", "1" = "Heterozygous")) +
  scale_x_discrete(labels=c("1" = "Male", "2" = "Female")) +
  scale_y_continuous(breaks = seq(0, 100, by = 10))
ggb
ggsave(paste("PLOT/sexage.pdf", sep = ""), height = 4 , width = 4)
ggsave(paste("PLOT/sexage.png", sep = ""), height = 4 , width = 4)
```
### General statistics
```{r}
library(table1)
summary(dat_total)
dat_total %>%
  group_by(genotype) %>%
  table1(age, sex, birth_year)

summary(dat_total)
dat_total %>%
  group_by(genotype) %>%
  table1(age, sex, birth_year, death_age)
```

